Worldwide estimation of parameters for a simple reaction-diffusion model of urban growth

### 2019-12-04 Docs/Communications/2019/ILUS2019/Communication/ilus2019.tex
.
maps_est/phase-diags

\title[Models of urban morphogenesis]{Models of urban morphogenesis to link urban form and function}
\author[J. Raimbault]{J.~Raimbault$^{1,2,3\ast}$\\\medskip
$^{\ast}$\texttt{j.raimbault@ucl.ac.uk}
}

@@

\date[15th November 2019]{TQG Debates 2019\\
3.1: Fractals and Multi-fractals\\
November 15th 2019
}

@@

%  Understanding drivers of urban growth and land-use change at the metropolitan scale often requires to isolate and estimate stylized processes of urban evolution. Several land-use change and urban growth models for example based on cellular automata have been introduced in the literature, but their lack of generality and need for rich data may become a shortcoming. We propose here to apply and calibrate a simple model of urban growth based on reaction-diffusion processes for population density, providing a rather general insight into world urban areas from simple data. This four-parameter model introduced by \cite{raimbault2018calibration} combines aggregation forces (positive externalities) with dispersal forces (negative externalities) to capture dynamics of urban form. The model was previously only simulated on synthetic initially empty spaces and statically calibrated by comparing the projection of final simulated configurations in a space of urban morphology indicators (Moran index, entropy, average distance, hierarchy) to real measures computed on a fixed-size moving window for Europe. We significantly extend these result by (i) applying the model on the 1000 largest urban areas of the Global Human Settlements database \citep{Florczyk2019ghs}; (ii) proceeding to a calibration on three successive time windows (1975-1990, 1990-2000 and 2000-2015) what provides dynamical values for parameters; (iii) using genetic optimization algorithms to minimise the distance on morphological indicators for each area. The computational cost being high, results are obtained by using the OpenMOLE model exploration software \citep{reuillon2013openmole}, which provides in a transparent way the optimization algorithms and the distribution of calibrations on a computation grid. We obtain therein comparable values for aggregation, diffusion and growth speed parameters on all these areas for the three periods given above. These values can be used for short terms projections, but also high-level policies by situating urban areas in the model phase diagrams and for example comparing them to more or less close phase transition parameter values. We also illustrate possible policy applications by relating estimated parameters to the EDGAR greenhouse gases emission database. This work thus shows how simple models can be useful for global comparisons, which are necessary to contextualize territorial policies.

@@

% center of Paris
%\includegraphics[width=0.88\textwidth]{figures/intro_paname}
%\includegraphics[width=0.9\textwidth]{figures/intro_bp}

@@

@@

\sframe{Which models for Urban Morphogenesis ?}{

% why modeling, exemple of rbd
% -> champ ; difficultés ("verrous". rq : j'aime pas la metaphore du verrou, trop cadrant et suppose que deja construit et qu'il suffit d'ouvrir. demarche epistemo de co-evol des connaissances, ni inductif ni deductif. il faut contruire. horizons d'exploration plus appropriés. rejoint l'idee de faire rentrer dans "cadre analytique" : cases prédéfinies, alors qu'il s'agit au contraire de définir ces cases. -> relire hdr arnaud

\begin{columns}
\column{0.35\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/intro_RBD_lattice}\\
\footnotesize\textit{Example: a basic hybrid model based on elementary processes for density and network \cite{raimbault2014hybrid}}
\column{0.6\textwidth}
\justify

%\vspace{-1cm}

\end{columns}

\textbf{Research Objective : } Explore simple models to capture morphogenesis based on abstract representation of urban processes; test their ability to reproduce existing urban systems.

@@

% comment Arnaud : reaction-diffusion ?

% model rationale and processes

@@

\sframe{Model Formalization}{

% model formalization and indicators

@@

\begin{enumerate}
\item Population growth with exogenous rate $N_G$, attributed independently to a cell following a preferential attachment of strength $\alpha$
%\begin{equation}
%\Pb{P_i(t+1)=P_i(t)+1|P(t+1)=P(t)+1}=\frac{(P_i(t)/P(t))^{\alpha}}{\sum(P_i(t)/P(t))^{\alpha}}
%\end{equation}
%The attribution being uniformly drawn if all population are equal to 0.
\item Population is diffused $n_d$ times with strength $\beta$
\end{enumerate}

@@

@@

@@

\sframe{Empirical Data for Calibration}{

% comments Arnaud : Quel lien avec les slides avant et après ? "Real" : le terme me paraît discutable : données empiriques plutôt ?
%  Décrire les données car ici tu passes directement aux indicateurs

\sframe{Model Calibration}{

@@

@@

@@

\section{Co-evolution model}

\sframe{Including more complex processes ?}{

% transition : representation of territories

\textit{Which ontology to include more complex functional properties ?}

@@

$\rightarrow$ Territorial systems as the strong coupling between territories and (potential and realized) networks \cite{dupuy1987vers}.

@@

$\rightarrow$ Networks convey functional notions of centralities and accessibility, among others; have furthermore proper topological properties.

@@

@@

@@

\sframe{Morphological indicators}{

>>>>

\title[]{Worldwide estimation of parameters for a simple reaction-diffusion model of urban growth}
\author[J. Raimbault]{J.~Raimbault$^{1,2,3\ast}$\\\medskip
$^{\ast}$\texttt{j.raimbault@ucl.ac.uk}
}

@@

\date[5th December 2019]{International Land-Use Symposium 2019\\
3.1: Fractals and Multi-fractals\\
5th December 2019
}

@@

@@

\includegraphics[width=0.88\textwidth]{figures/intro_paname}

@@

\includegraphics[width=0.9\textwidth]{figures/intro_bp}

\sframe{Modeling urban growth and land-use change}{

%  Understanding drivers of urban growth and land-use change at the metropolitan scale often requires to isolate and estimate stylized processes of urban evolution. Several land-use change and urban growth models for example based on cellular automata have been introduced in the literature, but their lack of generality and need for rich data may become a shortcoming.

\cite{engelen2008validating} calibration/validation + map comparison

\cite{zhou2018gini} urban form / emissions

@@

\sframe{Contribution}{

%  We propose here to apply and calibrate a simple model of urban growth based on reaction-diffusion processes for population density, providing a rather general insight into world urban areas from simple data.

% We significantly extend these result by (i) applying the model on the 1000 largest urban areas of the Global Human Settlements database \citep{Florczyk2019ghs}; (ii) proceeding to a calibration on three successive time windows (1975-1990, 1990-2000 and 2000-2015) what provides dynamical values for parameters; (iii) using genetic optimization algorithms to minimise the distance on morphological indicators for each area.

Explore simple models to capture morphogenesis based on abstract representation of urban processes; test their ability to reproduce existing urban systems.

@@

% This four-parameter model introduced by \cite{raimbault2018calibration} combines aggregation forces (positive externalities) with dispersal forces (negative externalities) to capture dynamics of urban form. The model was previously only simulated on synthetic initially empty spaces and statically calibrated by comparing the projection of final simulated configurations in a space of urban morphology indicators (Moran index, entropy, average distance, hierarchy) to real measures computed on a fixed-size moving window for Europe.

@@

\sframe{Model formalization}{

@@

\begin{enumerate}
\item Population growth with exogenous rate $N_G$, attributed independently to a cell following a preferential attachment of strength $\alpha$
\item Population is diffused $n_d$ times with strength $\beta$
\end{enumerate}

@@

\sframe{Measuring urban morphology}{

@@

\sframe{Model behavior (synthetic data))}{

@@

\sframe{Model application on a spatial grid}{

@@

\section{Empirical analysis}

\sframe{GHSL database}{
	% brief description of GHSL; illustration

% + empirical approach: 1000 areas, window 0.25%

$\rightarrow$ Application on worldwide urban areas using the GHSL population grid

$\rightarrow$ Dynamical calibration on three successive time windows (1975-1990, 1990-2000, 2000-2015)

$\rightarrow$ Extraction of neighborhood (spatial span times 1.5) of 1000 larger urban areas (2015 population)

$\rightarrow$ Computation of morphological indicators on corresponding population grids

\sframe{Global indicators distribution: Moran}{

@@

\includegraphics[width=\linewidth]{figures/mapindic_moran_2015.png}

\small

\textit{Regional patterns for Moran index}

@@

\sframe{Global indicators distribution: average distance}{

\includegraphics[width=\linewidth]{figures/mapindic_avgDist_2015.png}

@@

\small

\textit{Complementarity of average distance indicator}

\sframe{Clustering morphological trajectories}{

\includegraphics[width=0.55\textwidth]{figures/distributions_cluster-year.png}
\includegraphics[width=0.4\textwidth]{figures/PCA_trajectories.png}

@@

\small

\textit{Disjoint distributions and temporal trajectories: model application is relevant}

\sframe{Distribution of clusters}{

\includegraphics[width=\linewidth]{figures/mapindic_cluster.png}

\sframe{Correlations}{

\includegraphics[width=0.48\linewidth]{figures/correlations_deltavars-morpho_2015.png}
\includegraphics[width=0.48\linewidth]{figures/correlations_vars-morpho_2015.png}

\section{Model calibration}

\sframe{Calibration method}{
 % setup, error
}

\sframe{Implementation}{

% openmole slide

% The computational cost being high, results are obtained by using the OpenMOLE model exploration software \citep{reuillon2013openmole}, which provides in a transparent way the optimization algorithms and the distribution of calibrations on a computation grid.

%\footnotesize
\justify

\textit{Performance constraints: simulate $N$ mesoscopic morphogenesis models in parallel (macroscopic interactions are efficient as based on matrices)}

$\rightarrow$ model implemented in \texttt{scala} and integrated within a broader library (including implementations of \cite{raimbault2018indirect} \cite{raimbault2018calibration}

\cite{favaro2011gibrat} \cite{cottineau2015modular})

\textit{Large number of parameters and output indicators}

$\rightarrow$ integration into the OpenMOLE model exploration open source software \cite{reuillon2013openmole}

\begin{center}
\includegraphics[height=0.13\textheight]{figures/iconOM.png}
\includegraphics[height=0.13\textheight]{figures/openmole.png}
\end{center}

%(X year of computation for the results presented here)

\footnotesize
\textit{Enables seamlessly (i) model embedding; (ii) access to HPC resources; (iii) exploration and optimization algorithms}

\textbf{Come to the satellite \textit{New Methods and Epistemologies to Explore Simulation Models} tomorrow afternoon in LHN-TR+05}

\sframe{Examples of results}{

% pareto fronts for two different areas

\includegraphics[width=0.49\textwidth]{figures/pareto_indics-relpop_1_coloralpha.png}
\includegraphics[width=0.49\textwidth]{figures/pareto_indics-relpop_100_coloralpha.png}

\sframe{Adjusted parameters}{

% maps : alpha, beta, tstep

% We obtain therein comparable values for aggregation, diffusion and growth speed parameters on all these areas for the three periods given above.

\includegraphics[height=0.49\textheight]{figures/mapparam_meanalpha_2015.png}
\includegraphics[height=0.49\textheight]{figures/mapparam_meanbeta_2015.png}

\sframe{Trajectories in phase diagrams}{

% These values can be used for short terms projections, but also high-level policies by situating urban areas in the model phase diagrams and for example comparing them to more or less close phase transition parameter values.

\sframe{Correlations}{

% We also illustrate possible policy applications by relating estimated parameters to the EDGAR greenhouse gases emission database.

\includegraphics[width=0.49\textwidth]{figures/deltavar-params_2000.png}
\includegraphics[width=0.49\textwidth]{figures/deltavar-params_2015.png}

@@

%  This work thus shows how simple models can be useful for global comparisons, which are necessary to contextualize territorial policies.

% \textit{Which ontology to include more complex functional properties ?}
%$\rightarrow$ Territorial systems as the strong coupling between territories and (potential and realized) networks \cite{dupuy1987vers}.
%$\rightarrow$ Networks convey functional notions of centralities and accessibility, among others; have furthermore proper topological properties.

@@

\sframe{Empirical Data (moving window)}{

@@
